---
title: "Draft: Methods"
---

```{r Packages}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(maps)
library(ggplot2)
```

- introduction
- clustering
  - creating "groups" of the data without knowing which one is which
  - seeing where the Francis Scott key bridge falls within the clusters
- KNN
  - simple ranking by distance with specific variables, excluding place, state code, and other location indicator variables
  - we will choose the top 5 "closest" bridges
  - talk about them individually in the results section

```{r Data Wrangling}
#| code-fold: true

bridges = read.csv("NTAD_National_Bridge_Inventory_5655614151403109778.csv")

ne_bridges = bridges %>% 
  filter(STATE_CODE_001 == 31, DECK_COND_058 != "N",
         SUBSTRUCTURE_COND_060 != "N", SUPERSTRUCTURE_COND_059 != "N",
         SERVICE_UND_042B %in% c("5","6","7","8","9")) %>%
  select(RECORD_TYPE_005A,YEAR_BUILT_027,TRAFFIC_LANES_ON_028A,
         TRAFFIC_LANES_UND_028B,ADT_029,YEAR_ADT_030,SERVICE_UND_042B,
         MAX_SPAN_LEN_MT_048,STRUCTURE_LEN_MT_049,ROADWAY_WIDTH_MT_051,
         DECK_WIDTH_MT_052,VERT_CLR_OVER_MT_053,DECK_COND_058,
         SUPERSTRUCTURE_COND_059,SUBSTRUCTURE_COND_060,LATDD,LONGDD)
```

```{r Awesome State Graph}
#| code-fold: true

nebraska = map_data("state") %>%
  filter(region == "nebraska")

ggplot() +
  geom_polygon(data = nebraska,aes(x = long, y = lat, group = group),
               fill = "white", color = "black", linewidth = 1) +
  geom_point(data = ne_bridges,aes(x = LONGDD, y = LATDD),
    color = "darkblue", size = 1) +
  coord_fixed(1.3) +
  theme_minimal() +
  labs(x = "Longitude",y = "Latitude",title = "Bridges over Waterways in Nebraska")
```